Relative parabolic factorizations of Chevalley groups

### 2015-05-22 bass-kolster.tex
Minor fixes

\noindent Then the principal congruence subgroup $\G(\Phi,R,I)$ admits the following relative version of Bass---Kolster decomposition:
\[\G(\Phi,R,I)=  \G(\Delta_1,R,I)\cdot \U(\Sigma_1,I)\cdot \U(\Sigma_1^-\setminus\{-\alpha_\mathrm{max}\},I)\cdot\Z \cdot \U(\Phi^-,I)\cdot \U(\Phi^+,I), \]
where $\Z=\left\{ z_{-\alpha_\mathrm{max}}(r,1)\mid r\in I \right\}$.
\end{thm}
\begin{proof}

@@

\textsc{Case $\Phi=\rC_\ell$, $n=2\ell$:}
Notice that column $(v_1,\ldots, v_{-2}, v_{-1}^2)^t$ is also $I$-unimodular.
Applying condition $\sr(I)\leq 2\ell-1$ we find $c_1, c_2, \ldots, c_{-2} \in I \cdot v_{-1}$ such that upper $2\ell -1$ components of $v'=(v_1 + c_1 v_{-1}, \ldots, v_{-2} + c_{-2}v_{-1}, v_{-1})^t$ form an $I$-unimodular column.
By the choice of $c_i$ we can find suitable $d\in I$ such that $h_1 \cdot v = v'$ for $h_1 = t_{1,-1}(c_1 + d) \cdot \prod_{i=2}^{-2} t_{i,-1}(c_i) \in \U(\Sigma_1, I)$.

@@

Using standard elementary calculations we can decrease the number of elementary root unipotents involved in the decomposition of the previous theorem.

\begin{cor}
In the assumptions and notation of Theorem~\ref{thm:BassKolster} one can write
\[\G(\Phi,R,I)=  \G(\Delta_1,R,I)\cdot \U(\Sigma_1,I)\cdot \U(\Sigma_1^-,I)\cdot\Z \cdot \U(\Sigma^-_1,I)\cdot \U(\Sigma_1, I).\]
\end{cor}

>>>>

\noindent Then the principal congruence subgroup $\G(\Phi,R,I)$ admits the following relative version of Bass---Kolster decomposition:
\[\G(\Phi,R,I)=  \U(\Phi^+,I) \cdot \U(\Phi^-,I) \cdot Z \cdot \U(\Sigma_1^-\setminus\{-\alpha_\mathrm{max}\},I) \cdot \U(\Sigma_1,I) \cdot \G(\Delta_1,R,I)\]
where $\Z=\left\{ z_{-\alpha_\mathrm{max}}(r,1)\mid r\in I \right\}$.
\end{thm}
\begin{proof}

@@

\textsc{Case $\Phi=\rC_\ell$, $n=2\ell$:}
Notice that column $(v_1,\ldots, v_{-2}, v_{-1}^2)^t$ is also $I$-unimodular.
Applying condition $\sr(I)\leq 2\ell-1$ we find $c_1, c_2, \ldots, c_{-2} \in I \cdot v_{-1}$ such that upper $2\ell -1$ components of $v'=(v_1 + c_1 v_{-1}, \ldots, v_{-2} + c_{-2}v_{-1}, v_{-1})^t$ form an $I$-unimodular column.
By the choice of $c_i$ we can find suitable $d\in I$ such that $h_1 \cdot v = v'$ for $h_1 = t_{1,-1}(c_1 + d) \cdot \prod_{i=2}^{-2} t_{i,-1}(c_i) \in \U(\Sigma_1^-, I)$.

@@

We can estimate the number of elementary root unipotents involved in the decomposition of the previous theorem.
\begin{lemma}
In the assumptions and notation of Theorem~\ref{thm:BassKolster} one can write any element $g$ of 
$\G(\Phi,R,I)$ as a product of an element of $\G(\Delta_1,R,I)$, element $z_{-\alpha}(s, 1)$ for some $s\in I$ and $n(\Phi)$ elementary unipotents $t_{\alpha_i}(s_i)$, where
\begin{itemize}
 \item $n(\rA_\ell)= 4 \ell -1$;
 \item $n(\rC_\ell)= 8 \ell - 5$;
 \item $n(\rD_\ell)= $
\end{itemize}

\cdot \U(\Sigma_1,I)\cdot \U(\Sigma_1^-,I)\cdot\Z \cdot \U(\Sigma^-_1,I)\cdot \U(\Sigma_1, I).\]
\end{lemma}
